---
title: "Assignment 3"
author: "Wang Pei"
date: "4/3/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r loading packages, include=FALSE}
library(dplyr)
library(car)
library(lfe)
library(stargazer)
library(evd) # gumbel distribution
library(stats4) # mle
library(AER)
```

# Question 1
### The sales department plans to run a large promotion. They provide you data obtained from a small trial (“hw3q1.csv”). They want to understand the effectiveness of the promotion. (Hint: the effectiveness can be different for different types of consumers.)

```{r load_data}
df1 <- read_csv("hw3q1.csv")
df1$age <- as.factor(df1$age)  ## Observe 3 age categories: 20-30, 30-40, 40+
df1$received_coupon <- as.factor(df1$received_coupon)
df1$after <- as.factor(df1$after)
summary(df1)

df1_2030 <- df1 %>% filter(age == "20-30")
df1_3040 <- df1 %>% filter(age == "30-40")
df1_40 <- df1 %>% filter(age == "40+")

nrow(df1_2030[df1_2030$after==1,])/nrow(df1_2030)  # 50% each before and after
nrow(df1_3040[df1_3040$after==1,])/nrow(df1_3040)  # 50% each before and after
nrow(df1_40[df1_40$after==1,])/nrow(df1_40)  # 50% each for before and after

nrow(df1_2030[df1_2030$received_coupon==1,])/nrow(df1_2030)  # 50% each treatment and control
nrow(df1_3040[df1_3040$received_coupon==1,])/nrow(df1_3040)  # 50% each treatment and control
nrow(df1_40[df1_40$received_coupon==1,])/nrow(df1_40)  # 50% each for treatment and control
```

## Part I 
### Calculate the treatment effects. Verify your results using felm. Are they the same as before-vs-after comparisons or treated-vs-non-treated comparisons?

Calculate the treatment effects using difference-in-differences (DID) estimator:
Treatment effect = (Y_treated, after - Y_nontreated, after) - (Y_treated, before - Y_nontreated, before)

```{r DID}
# Treatment effect for all types of customers
overall_diff <- df1 %>% group_by(received_coupon, after) %>% summarise(rev_mean = mean(revenue))
overall_diff

# Treatment effect for all types of customers
(11.549727 - 10.535157) - (9.493847 - 9.348597)   # 0.86932

# Treatment effect for difference types of customers (breakdown)
sep_diff <- df1 %>% group_by(age, received_coupon, after) %>% summarise(rev_mean = mean(revenue))
sep_diff

# Treatment effect for age group 20-30 using DID: 
(12.32244 - 11.78103) - (10.36813 - 10.31286)   # 0.48614

# Treatment effect for age group 30-40 using DID:
(10.32048 - 9.10888) - (8.77142 - 8.45867)   # 0.89885

# Treatment effect for age group 40+ using DID: 
(12.00626 - 10.71556) - (9.34199 - 9.27426)   # 1.22297
```

Answer: The average treatment effect for all types of customers i.e. across all age groups is 0.86932. The treatment effects differ for customers from different age groups. For customers aged from 20 to 30, the treatment effect is 0.48614. For customers aged from 30 to 40, the treatment effect is 0.89885. For customers aged 40 and above, the treatment effect is 1.22297.



Verify the results using felm:
felm can be used to verify the results of average treatment effect obtained using before-vs-after comparisons and treated-vs-nontreated comparisons:

1. felm to verify results of before-vs-after comparisons
```{r felm}
lm1 <- felm(revenue ~ after | received_coupon, data = df1_2030)
lm2 <- felm(revenue ~ after | received_coupon, data = df1_3040)
lm3 <- felm(revenue ~ after | received_coupon, data = df1_40)
stargazer(lm1, lm2, lm3, type="text",omit.stat=c("LL","ser","f"), model.numbers=TRUE)
```
For customers aged from 20 to 30, the coefficient of the binary variable after is 1.711.
For customers aged from 30 to 40, the coefficient of the binary variable after is 1.100.
For customers aged 40 and above, the coefficient of the binary variable after is 2.053.

Using before-vs-after comparisons: avergae of Y_after - average of Y_before
```{r before-after}
# Age 20-30
(12.32244 + 11.78103) / 2 - (10.36813 + 10.31286) / 2   # 1.71124

# Age 30-40
(10.32048 + 9.10888) / 2 - (8.77142 + 8.45867) / 2   # 1.099635

# Age 40+
(12.00626 + 10.71556) / 2 - (9.34199 + 9.27426) / 2   # 2.052785
```
For customers aged from 20 to 30, the before-vs-after comparison is 1.71124.
For customers aged from 30 to 40, the before-vs-after comparison is 1.099635.
For customers aged 40 and above, the before-vs-after comparison is 2.052785.

Thus, the results of felm which makes treatment (received_coupon) a fixed effect are consistent with the results of before-vs-after comparisons for different age groups.


2. felm to verify results of treated-vs-nontreated comparisons
```{r felm}
lm4 <- felm(revenue ~ received_coupon | after, data = df1_2030)
lm5 <- felm(revenue ~ received_coupon | after, data = df1_3040)
lm6 <- felm(revenue ~ received_coupon | after, data = df1_40)
stargazer(lm4, lm5, lm6, type="text",omit.stat=c("LL","ser","f"), model.numbers=TRUE)
```
For customers aged from 20 to 30, the coefficient of the binary variable received_coupon is 0.298.
For customers aged from 30 to 40, the coefficient of the binary variable received_coupon is 0.762.
For customers aged 40 and above, the coefficient of the binary variable received_coupon is 0.679.

Using treated-vs-nontreated comparisons: avergae of Y_treated - average of Y_nontreated
```{r treated-nontreated}
# Age 20-30
(10.36813 + 12.32244) / 2 - (10.31286 + 11.78103) / 2   # 0.29834

# Age 30-40
(8.77142 + 10.32048) / 2 - (8.45867 + 9.10888) / 2   # 0.762175

# Age 40+
(9.34199 + 12.00626) / 2 - (9.27426 + 10.71556) / 2   # 0.679215
```
For customers aged from 20 to 30, the treated-vs-nontreated comparison is 0.29834.
For customers aged from 30 to 40, the treated-vs-nontreated comparison is 0.762175.
For customers aged 40 and above, the treated-vs-nontreated comparison is 0.679215.

Thus, the results of felm which makes time (after) a fixed effect are consistent with the results of treated-vs-nontreated comparisons for different age groups.


felm can verify before-vs-after comparisons and treated-vs-nontreated comparisons, but it cannot verify the treatement effects computed using difference-in-differences (DID) estimator. This is because felm fits a linear model by controling for fixed effects but DID does not require to fix any variation.
- In the case of using felm to verify before-vs-after comparisons:
felm makes treatment (received_coupon) a fixed effect. received_coupon controls for any unobserved treatment-specific heterogeneity that is is persistent over the time of experiment. Thus, there is no time trend in the model by setting received_coupon a fixed effect. This satisifies the assumption of before-vs-after comparisons. The felm results are consistent with before-vs-after comparisons.
- In the case of using felm to verify treated-vs-nontreated comparisons:
felm makes time (after) a fixed effect. after controls for any unobserved time-specific trend that is common across all the customers. Thus, each customers have the same probability of being chosen as treatment and the experiment can be assumed to be random assignment of treatment. This satisifies the assumption of treated-vs-nontreated comparisons. The felm results are consistent with treated-vs-nontreated comparisons.

However, difference-in-differences does not require any of the above assumptions. The two assumption are violated in the sample data: there is a time trend and the assignment of treatment is not random. Thus, treatment effects measured using DID estimators are different from both before-vs-after comparisons and treated-vs-nontreated comparisons. felm is unable to verfiy the treatment effects computed using DID estimators.


## Part II 
### The cost of the promotion is $0.8 per user. Provide specific advices to your colleague and estimate the gain (if any) from the promotion.

From Part I:
The average treatment effect for age group 20-30 is 0.48614, which is the revenue gain per user in the age group of 20-30.
The average treatment effect for age group 30-40 is 0.89885, which is the revenue gain per user in the age group of 30-40.
The average treatment effect for age group 40+ is 1.22297, which is the revenue gain per user in the age group of 40+.

Given that the cost of the promotion is $0.8 per user, below are the gain from the promotion for each age group of customers and some findings:
(1) It is profitable to run this promotion among age groups 30-40 and 40+, and 40+ is the most profitable age group.
Profit gain per user for age group 30-40 = 0.89885 - 0.8 = 0.09885
Profit gain per user for age group 40+ = 1.22297 - 0.8 = 0.42297
2) It is unprofitable to run this promotion among age group 20-30.
Profit loss per user for age group 20-30 = 0.48614 - 0.8 = -0.31386
3) Overall gain from the promotion based on the small trial of 12000 by combining 3 age groups is 
4000 * 0.09885 + 4000 * 0.42297 + 4000 * (-0.31386) = 831.84

Here are some advices to our colleagues:
(1) It is favourable to run the promotion among age groups of 30-40 and 40+, and the company should promote the coverage of the promotion especially among those aged 40+ i.e. reach out more customers aged 30-40 and 40+, issue more vouchers to them, and encourage them to particpate in the promotion.
(2) It is unfavourable to run the promotion for age group 20-30, and the company should reduce the coverage of the promotion among those aged 20-30 i.e. issue less vouchers to those aged 20-30.
(3) The company should issue vouchers based on customers' age. This can be achieved by showing ID to qualify for claiming the voucher or use internal data to filter the targeted customers and only issue vouchers to those aged 30-40 and 40+.


# Q2

```{r Q2}
data2 = read.csv("hw3q2.csv")
```

```{r Q2.1 mle}
lik = function(beta0,beta1,beta2) {
  # likelihood function
  Pij =
  (data2$choices==0) * 0 +
	(data2$choices==1) * (beta1 + data2$p1 * beta0) +
	(data2$choices==2) * (beta2 + data2$p2 * beta0) -
	log(1+exp(beta1 + data2$p1*beta0)+exp(beta2 + data2$p2*beta0))
  
	return(-sum(Pij))
}

# estimation
mle1 = mle(lik, 
	start = runif(3,-1,0),  # random initialization
	method = "L-BFGS-B", 
	lower = c(-10, -10,-10),
  upper = c(-0.00001, 10, 10))  # beta0, coeff of price, should be negtive

summary(mle1)
#  beta0       beta1        beta2
# -0.07906284  0.66639176  -1.66095162
```
Interpretation of coefficients:

- beta0 = -0.07906249
Price coefficient: One unit increase in price of a product, either JustGrab or GrabShare, will result in a 0.0791 units decrease in the utility level in average.

- beta1 = 0.66638432
The preference for JustGrab (j = 1) relative to not choosing any services is higher i.e. JustGrab is a more attractive option / generates higher utility than not choosing any (outside option).

- beta2 = -1.66094764
The preference for GrabShare (j = 2) relative to not choosing any services is lower i.e. GrabShare is a less attractive option / generates lower utility than not choosing any (outside option).


## Q2.2 
Endogeneity issue:
Demand for GrabShare is unobserved, being an omitted variable which results in the endogeneity issue. During the morning peak hours when many people rush for school and work, the demand for GrabShare is high. This increase in demand leads to price surge, which results in higher price. Thus, demand for GrabShare is correlated with price2. Also, higher demand implies higher utility gained by customer once they obtain the GrabShare service, which affects customer choice. This suggests demand affects choice. Therefore, omitting the demand variable in the model will lead to omitted vairable bias, resulting in endogeneity.

Solution:
Demand for GrabShare is reflected by surge2 which is demand-driven. Because the underlying demand is unobserved to us, we can include surge2 in the model instead. The solution is to use surge as an intrumental variable. Surge pricing only affects customer's utility level / choice only through the final price. Therefore, surge price is a good intrumental variable when running regression of utility on price.

```{r Q2.2}
data2$sales = 1

mkt_share=aggregate(sales~p1+p2+surge_1+surge_2+choices, data=data2, FUN=sum)
mkt_share$total=ave(mkt_share$sales,by=list(mkt_share$p1,mkt_share$p2,mkt_share$surge_1,mkt_share$surge_2),FUN=sum)
mkt_share$share=mkt_share$sales/mkt_share$total


iop_dta=subset(mkt_share,choices!=0)#inside options
oop_dta=subset(mkt_share,choices==0)[,c("p1","p2","share")]#outside option
colnames(oop_dta)=c("p1","p2","share0")
lm_dta=merge(oop_dta,iop_dta)#merge inside options with outside option

lm_dta$log_odds=log(lm_dta$share/(lm_dta$share0))
lm_dta$price=(lm_dta$choices==1)*lm_dta$p1+
(lm_dta$choices==2)*lm_dta$p2   #the price for the focal product
lm_dta$product=as.factor(as.character(lm_dta$choices))
lm_dta[1:20,]

lm_dta$surge = (lm_dta$choices==1)*(lm_dta$surge_1) + (lm_dta$choices==2)*(lm_dta$surge_2)

logistic_regression=lm(log_odds~price+product,lm_dta)
iv_reg=ivreg(log_odds ~ price+product|surge+product, data=lm_dta)
stargazer(logistic_regression,iv_reg, type="text",omit.stat=c("f"))
```

beta0 = -0.213
beta1 = 1.685
beta2 = 1.685 - 2.557

```{r Q2.3.1 one product}
beta0 = -0.213
beta1 = 1.685
beta2 = 1.685 - 2.557
# JustGrab only
profit_function1 = function(price1) {
  Profit1 = price1 * (exp(beta1 + price1 * beta0) / (1 + exp(beta1 + price1 * beta0)))

	return(-Profit1)
}

# GrabShare only
profit_function2 = function(price2) {
  Profit2 = price2 * (exp(beta2 + price2 * beta0) / (1 + exp(beta2 + price2 * beta0)))

	return(-Profit2)
}

# Prices
p1 = optim(c(1), profit_function1, method = "BFGS")$par
p2 = optim(c(1), profit_function2, method = "BFGS")$par
c(p1, p2)
# Profits
c(-profit_function1(p1), -profit_function2(p2))
```


```{r Q2.3.2 two products}
beta0 = -0.213
beta1 = 1.685
beta2 = 1.685 - 2.557

profit_function3=function(prices){
  price1 = prices[1]
	price2 = prices[2]

	Profit1 = price1 * (exp(beta1 + price1 * beta0) / (1 + exp(beta1 + price1 * beta0) + exp(beta2 + price2 * beta0)))

	Profit2 = price2 * (exp(beta2 + price2 * beta0) / (1 + exp(beta1 + price1 * beta0) + exp(beta2 + price2 * beta0)))

	return(-(Profit1+Profit2))
}

res = optim(c(1, 1), profit_function3, method = "BFGS")$par
p1 = res[1]
p2 = res[2]

# Prices
c(p1,p2)
# Profit
-profit_function3(res)
```



```{r Q3.1.1 pre-merger}
set.seed(37)
#parameters
beta0=-0.03
beta1=2
beta2=1
beta3=2
N = 10
# global variables
price1=10
price2=10
price3=10

prob_1 = function(price1) {
  exp(beta1+price1*beta0)/
	(1+exp(beta1+price1*beta0)+exp(beta2+price2*beta0)+exp(beta3+price3*beta0))
}
prob_2 = function(price2) {
  exp(beta2+price2*beta0)/
	(1+exp(beta1+price1*beta0)+exp(beta2+price2*beta0)+exp(beta3+price3*beta0))
}
prob_3 = function(price3) {
  exp(beta3+price3*beta0)/
	(1+exp(beta1+price1*beta0)+exp(beta2+price2*beta0)+exp(beta3+price3*beta0))
}

cost = function(s) {
  (s-7)^2 + 50
}

# objective function  of firm 1
profit_1=function(price1){
  produce1 = N*prob_1(price1)
	profit = produce1*price1 - cost(produce1)
	return(-profit)
}

# objective function  of firm 2
profit_2=function(price2){
  produce2 = N*prob_2(price2)
	profit = produce2*price2 - cost(produce2)
	return(-profit)
}

# objective function  of firm 3
profit_3=function(price3){
  produce3 = N*prob_3(price3)
	profit = produce3*price3 - cost(produce3)
	return(-profit)
}

# best response dynamic 
i=0
error=1
Iter=1000
Tol=1e-10
prices = c(1,1,1)
while (i<Iter & error>Tol) {

	i=i+1
	price1=prices[1]
	price2=prices[2]
	price3=prices[3]
	
	prices[1] = optim(c(1), profit_1, method = "L-BFGS-B",
	             lower=c(0),upper=c(Inf))$par
	
	prices[2] = optim(c(1), profit_2, method = "L-BFGS-B", 
	             lower=c(0),upper=c(Inf))$par
	
  prices[3] = optim(c(1), profit_3, method = "L-BFGS-B", 
	             lower=c(0),upper=c(Inf))$par
	
	error=abs(prices[1]-price1)+abs(prices[2]-price2)+abs(prices[3]-price3)
}

price1=prices[1]
price2=prices[2]
price3=prices[3]

c(i, error)
# prices
prices
# sales
c(N*prob_1(price1), N*prob_2(price2), N*prob_3(price3))
# costs
sapply(c(N*prob_1(price1), N*prob_2(price2), N*prob_3(price3)), cost)
# profits
c(-profit_1(c(price1)),-profit_2(c(price2)),-profit_3(c(price3)))

```
The prices of the three products before merger are 42.47629 30.04197 42.47629.
The sales of the three products before merger are 3.313246 1.769962 3.313246 (million).
The cost of the three products before merger is 63.59215 77.35330 63.59215 (million).
The profits of the three products before merger is 77.14225 -24.18015  77.14225.



```{r Q3.1.2 post-merger}
cost_merged = function(s1,s2,lambda1=1,lambda2=1) {
  return(lambda1*cost(s1) + lambda2*cost(s2))
}

profit_merged = function(param) {
  price1=param[1]
  price2=param[2]
  produce1=N*exp(beta1+price1*beta0)/
	(1+exp(beta1+price1*beta0)+exp(beta2+price2*beta0)+exp(beta3+price3*beta0))
  
  produce2=N*exp(beta2+price2*beta0)/
	(1+exp(beta1+price1*beta0)+exp(beta2+price2*beta0)+exp(beta3+price3*beta0))
  
  profit=produce1*price1 + produce2*price2 - cost_merged(produce1,produce2)
  
  return(-profit)
}

price3 = 42.47629
res = optim(c(1,1), profit_merged, method = "L-BFGS-B",
	             lower=c(0,0),upper=c(Inf,Inf))$par

price1 = res[1]
price2 = res[2]
price3 = 42.47629

# prices
c(price1, price2, price3)
# sales
c(N*prob_1(price1), N*prob_2(price2),N*prob_3(price3))
# costs
c(cost_merged(N*prob_1(price1), N*prob_2(price2)), cost(N*prob_3(price3)))
# profits
c(-profit_merged(c(price1, price1)), -profit_3(price3))

```
If firm 3’s price is held constant:

- The prices of the three products after merger are 50.38601 46.79862 42.47629.
- The sales of the three products after merger are 3.038511 1.244818 3.852256 (million).

- The cost of the post-merger firm and firm 3 after merger is 148.81552  59.90829 (million).
- The profits of the post-merger firm and firm 3 after merger is 52.9621 103.7213.

Compared with the pre-merger equilibrium,
the prices of product 1 and 2 increase; sales of product 1 and 2 decreases, sales of product 3 increases; the sum of cost of firm 1 and 2 increases from 63.59215+77.35330 = 140.9454 to 148.81552 and the cost of firm 3 decrease.
The profit of the merge increases from 77.14225-24.18015 = 52.9621 to 62.29782; the profit of firm 3 increases

```{r Q3.2.a}
i = 0
error=1
Iter=1000
Tol=1e-10
prices = c(1,1,1)

while (i < 150 & error > Tol) {
  price1 = prices[1]
  price2 = prices[2]
  price3 = prices[3]
  
  res1 = optim(c(1,1), profit_merged, method = "L-BFGS-B",
	             lower=c(0,0),upper=c(Inf,Inf))$par
  prices[1] = res1[1]
  prices[2] = res1[2]
  prices[3] = optim(c(1), profit_3, method = "L-BFGS-B",
	             lower=c(0),upper=c(Inf))$par
  
  i = i + 1
  error=abs(prices[1]-price1)+abs(prices[2]-price2)+abs(prices[3]-price3)
}

price1=prices[1]
price2=prices[2]
price3=prices[3]

c(i, error)
# prices
prices
# sales
c(N*prob_1(price1),N*prob_2(price2),N*prob_3(price3))
# costs
c(cost_merged(N*prob_1(price1),N*prob_2(price2)), cost(N*prob_3(price3)))
# profits
c(-profit_merged(c(price1, price2)),-profit_3(c(price3)))
```
If firm 3 responds as predicted by a simultaneous move pricing game:

- The prices of the three products after merger are 51.45059 47.79469 45.91185.
- The sales of the three products after merger are 3.100972 1.273021 3.661517(million).
- The cost of the merge and firm 3 after merger is 148.00072  61.14547 (million).
- The profits of the post-merger firm and firm 3 after merger is 72.38972 106.96156.

Compared with the answer in part I:
- The prices of product 1 and 2 increase and price of product 3 increases.
- The sales of product 1 and 2 increase and sales of price 3 decreases.
- The cost of the post-merger firm decreases and the cost of firm 3 increases.
- The profits of the post-merger firm increases and the profit of firm 3 decreases
```{r Q3.2.b}
merger_analysis = function(lambda1, lambda2) {
  cost_merged = function(s1,s2) {
    return(lambda1*cost(s1) + lambda2*cost(s2))
  }
  
  profit_merged = function(param) {
    price1=param[1]
    price2=param[2]
    produce1=N*exp(beta1+price1*beta0)/
  	(1+exp(beta1+price1*beta0)+exp(beta2+price2*beta0)+exp(beta3+price3*beta0))
    
    produce2=N*exp(beta2+price2*beta0)/
  	(1+exp(beta1+price1*beta0)+exp(beta2+price2*beta0)+exp(beta3+price3*beta0))
    
    profit=produce1*price1 + produce2*price2 - cost_merged(produce1,produce2)
    
    return(-profit)
  }
  
  profit_3=function(price3){
    produce3 = N*exp(beta3+price3*beta0)/
  	(1+exp(beta1+price1*beta0)+exp(beta2+price2*beta0)+exp(beta3+price3*beta0))
    
  	profit = produce3*price3 - cost(produce3)
  	return(-profit)
  }
  
  i = 0
  error=1
  Iter=1000
  Tol=1e-10
  prices = c(1,1,1)
  
  while (i < 150 & error > Tol) {
    price1 = prices[1]
    price2 = prices[2]
    price3 = prices[3]
    
    res1 = optim(c(1,1), profit_merged, method = "L-BFGS-B",
  	             lower=c(0,0),upper=c(Inf,Inf))$par
    prices[1] = res1[1]
    prices[2] = res1[2]
    prices[3] = optim(c(1), profit_3, method = "L-BFGS-B",
  	             lower=c(0),upper=c(Inf))$par
    
    i = i + 1
    error=abs(prices[1]-price1)+abs(prices[2]-price2)+abs(prices[3]-price3)
  }
  
  price1=prices[1]
  price2=prices[2]
  price3=prices[3]
  
  return (list(
    'prices'= prices, 
    'sales' = c(N*prob_1(price1),N*prob_2(price2),N*prob_3(price3)),
    'costs' = c(cost_merged(N*prob_1(price1),N*prob_2(price2)), cost(N*prob_3(price3))),
    'profits'=c(-profit_merged(c(price1, price2)),-profit_3(c(price3)))
  ))
  
}

#check its consistency with 3.2.a
merger_analysis(1,1)$profits[1]

lambdas = seq(0.5,1.5, by = 0.1)


record = rep(c(0,0,0), 121 )
record = array(record, dim = c(121,3))


row = 1
for (i in lambdas) {
  for (j in lambdas) {
    profit_aftermerge = merger_analysis(i,j)$profits[1]
    record[row,1] = i
    record[row,2] = j
    record[row,3] = profit_aftermerge
    row = row + 1
  }
  
}
#record
```


```{r Q3.2.2}
library("ggplot2")

record1 = as.data.frame(record)

names(record1) = c ("lambda1" , "lambda2", "profit_after_merge")

record1$merge = record1$profit_after_merge >  77.14225-24.18015
record1
ggplot(record1, aes(x=lambda1, y=lambda2)) + geom_point(aes(size=profit_after_merge, color = merge))
```
Profits for different values of \lambda_1 and \lambda_2 with the step of 0.1 from 0.5 to 1.5 are calculated. 
These profits are plotted in a scatter plot with x axis being the value of \lambda_1 and y axis being the value of \lambda_2. 
The size of the point represents the quantity of the merger's profit and the color represents whether the profit of firm 1 and 2 increases after merger. Blue means increase which indicates that under the \lambda_1 and \lambda_2 of a blue point, the merge should be conducted while red means decrease which indicates that under the \lambda_1 and \lambda_2 of a red point, the merge should not be conducted. 



```{r Q4 original params}
# set up
settings=list(K=3, # number of pack sizes
	n=c(2,6,0), # number of  units in size k, 0 is not buying
	Imax=20, # maximum inventory
	tol=1e-8, # error tolerance for convergence
	iter_max=10000  # maximum  number of iterations
	)

param=list(beta=0.99,
	alpha=4, # price sensitivity
	delta=10, # consumption utility
	c=0.05 # cost for holding inventory
	)

price=list(
	L=2, # number of price levels
	prices=data.frame(price_norm=c(2,5,0),price_prom=c(1.2,3,0)), #K-by-L, normal prices and promotion prices
	prob=c(0.84,0.16) # probability of different price levels.
	)

# fixed point iteration
value_function_iteration=function(settings,param,price){

	value_0=matrix(0,nrow=settings$Imax+1,ncol=price$L) # State vars are inventory and price
	error=settings$tol+1
	iteration=1
	start_time=Sys.time()
	
	while (error>=settings$tol & iteration<= settings$iter_max){
	  # bellman_operator takes in curr_value and updates State
		value=Bellman_operator(value_0,settings,param,price)$value		
		iteration=iteration+1
		error=max(abs(value-value_0))
		value_0=value
	}

	time_elapsed=Sys.time()-start_time
	solution=Bellman_operator(value_0,settings,param,price)

	output=list(time_elapsed,error,iteration,solution)
}

# fixed point operator that updates value function for any 
Bellman_operator=function(value_0,settings,param,price){
  
  # 3 dims: num_Inventory x num_Price x num_Choice
	v_choice=array(0,dim=c(settings$Imax+1,price$L,settings$K))
	value=matrix(0,nrow=settings$Imax+1,ncol=price$L)
	choice=matrix(0,nrow=settings$Imax+1,ncol=price$L)
	inventory=c(0:settings$Imax)
	Ev=value_0%*%price$prob  # expected value

	for (k in 1:(settings$K)){
	  # update inventory
		inventory_new0=inventory+settings$n[k]
		inventory_new=inventory_new0-1
		inventory_new[inventory_new<0]=0
		inventory_new[inventory_new>settings$Imax]=settings$Imax
		# inventory_new = min(settings$Imax, max(0, inventory_new0-1))
		
		# Calculate utility of curr inventory
		utility=param$delta*(inventory_new0>0)-param$c*inventory_new

		for (l in 1:price$L){
			v_choice[,l,k]=utility-param$alpha*price$prices[k,l]+param$beta*Ev[inventory_new+1]
		}		
	}

	for (i in 1:(settings$Imax+1)){
		for (l in 1:price$L){
			value[i,l]=max(v_choice[i,l,])
			choice[i,l]=which.max(v_choice[i,l,])
		}
	}

	output=list(value=value,choice=choice)
	return(output)
}

results=value_function_iteration(settings,param,price)
results
```


```{r Q4.1 Imax=5}
new_settings=list(K=3, # number of pack sizes
	n=c(2,6,0), # number of  units in size k, 0 is not buying
	Imax=5, # maximum inventory
	tol=1e-8, # error tolerance for convergence
	iter_max=10000  # maximum  number of iterations
	)

results=value_function_iteration(new_settings, param, price)
results
```
Describe and explain the findings:

- When maximum inventory capacity = 5, all maximum expected sum of discounted utility (abbreviated as utility value below) at each current inventory levels and price promotions are lower than those when maximum inventory capacity = 20.
  
- From this output, we can conclude that higher inventory capacity gives higher utility values. This is because the utility value of choice 2 (large package) is higher than that of choice 1 (small package) i.e. larger package is more value-for-money. When max inventory = 20, the customer never opt for choice 1, but when max inventory = 5, the customer seldom have chances to opt for choice 2.

- For both cases of normal price and promotion, the purchase choices tend to be 3 (not buying) when current inventory is close to maximum. This is because individuals with smaller inventory capacity are more likely to replenish the stock more frequently than those with larger inventory capacity under same rate of consumption. 

```{r Q4.2 discount=0.5}
new_param_1=list(beta=0.5,
	alpha=4, # price sensitivity
	delta=10, # consumption utility
	c=0.05 # cost for holding inventory
	)
results=value_function_iteration(settings, new_param_1, price)
results
```
Observation:

- When discounting factor = 0.5, all utility values at each current inventory levels and price promotions are lower than those when discounting factor = 0.99.

- The customer only buys product when current stock is 0, and he only opt for choice 1 (small package) which is cheaper (thus larger utility for today).

Explanation:

- The lower the discounting factor, the more the customer cares about today's utility.

- Based on policy function and Bellman equation, when the discounting factor gets smaller, the maximum expected sum of discounted utility gets smaller.

```{r Q4.3.1 price coeff larger}
new_param_2=list(beta=0.99,
	alpha=12, # price sensitivity: 4 -> 12
	delta=10, # consumption utility
	c=0.05 # cost for holding inventory
	)

new_param_3=list(beta=0.99,
	alpha=2, # price sensitivity: 4 -> 2
	delta=10, # consumption utility
	c=0.05 # cost for holding inventory
	)

results_1=value_function_iteration(settings, new_param_2, price)
results_2=value_function_iteration(settings, new_param_3, price)
c(results_1, results_2)
```
Observation:
- When price sensitivity is larger, all utility values at each current inventory levels and price promotions decrease.
- The customer only opt for choice 2 (buy large package) when current inventory is not close to full, and he always opt for choice 3 (not buying) when there is no promotions even though the current inventory is 0.

Explanation:

- When the price sensitivity is lower, the customer cares less about the price, thus the price dis-utility is lower. Based on the Bellman Operator, as a result, when the price dis-utility becomes less negative, the maximum expected sum of discounted utility increases.


```{r Q4.4}
new_price=list(
	L=3, # number of price levels
	prices=data.frame(price_norm=c(2,5,0),discount_1=c(1.4,3.5,0),discount_2=c(0.8, 2, 0)),
	# discount_1: 70%, discount_2: 40%
	#K-by-L, normal prices and promotion prices
	prob=c(0.8, 0.1, 0.1) # probability of different price levels.
	)

results=value_function_iteration(settings, param, new_price)
results

# Expected price for current settings
# Choice 1, small package
2*0.8 + 1.4*0.1 + 0.8*0.1
# Choice 2, large package
5*0.8 + 3.5*0.1 + 2*0.1

# Expected price for original settings
# Choice 1, small package
2*0.84 + 1.2*0.16
# Choice 2, large package
5*0.84 + 3*0.16
```

- After the price settings change, all utility values at each current inventory levels increase as compared to original settings.

- The expected prices for both choice 1 and choice 2 become lower than previous price settings: expected price for small package: 1.82 < 1.872, expected price for large package: 4.55 < 4.68. Thus, the lower the expected price, the less the price dis-utility, and the maximum expected sum of discounted utility is larger.

- When there is no promotion, the customer only opts for choice 1 (small package) when there is no stock; when there is a moderate promotion, the customer opts for choice 2 (large package) when the stock is low; when there is a great promotion, the customer opts for choice 2 (large package) as much as possible.

